% introduction.tex
% Introduction: motivation, technical gap, contribution, roadmap.

Portfolio construction, risk attribution, and stress testing often
take long synthetic time series of asset growth rates as inputs. A
good generator has to satisfy two things the downstream consumer
cares about. First, each single asset's synthetic series should look
like its real counterpart when you histogram it day by day: the tails
should be as heavy as real tails, the kurtosis should match, regimes
and jumps should show up with roughly the right frequency; in
statistical language this is the \emph{marginal distribution} of the
asset. Second, if we stack synthetic series across many assets they should
co-move the way real markets co-move: on a day when the market
plunges, stocks with high factor loadings should plunge with it, tail
co-movement should look realistic, and factor structure should be
intact; this is the \emph{cross-sectional dependence} across assets.
Theses two requirements are largely orthogonal, and off-the-shelf
generators tend to handle only one of them. Per-asset generators
built from hidden Markov models~\cite{hamilton1989}, generalized
autoregressive conditional heteroskedasticity (GARCH)
families~\cite{bollerslev1986}, or empirical
bootstraps~\cite{politisRomano1994} handle the first: fit one of
these models to AAPL's growth-rate history and its synthetic paths
carry AAPL's tail thickness, kurtosis, and regime structure. But the
fit knows nothing about market factors, e.g., SPY, or any other assets, e.g., MSFT. 
Stacking the per-asset draws therefore produces one synthetic
growth-rate path per asset, each matching its own asset's
full growth-rate marginal. But the paths carry no joint
structure: on a day when synthetic SPY plunges, synthetic AAPL has
no reason to plunge with it (but real AAPL would).

The standard recipe for adding factor structure is a linear factor
model, either the one-factor single-index model (SIM) of
Sharpe~\cite{sharpe1963} or a multi-factor variant such as Ross's
arbitrage pricing theory~\cite{rossAPT1976} or
Fama-French~\cite{famaFrench1993}. We develop the construction for
the one-factor SIM, which writes each asset's growth rate as an
intercept plus a market-loaded term plus an idiosyncratic residual;
the variance correction we propose extends to the multi-factor case
without structural change. Applied naively, the recipe plugs a
synthetic full growth-rate path from the per-asset generator (fit
to the asset's real growth-rate history and matching its full
growth-rate marginal) straight into the SIM as the idiosyncratic residual.
Because the full growth rate already contains the market-driven part,
using it as the residual double-counts the market variance and
inflates the marginal variance of the composed series by an amount
that grows quadratically in the factor loading. For high-loading
assets the inflation is large enough to be visible in the empirical
marginal: the body widens, the kurtosis drops, and the heavy tails
that the generator was built to deliver are pushed out into a
Gaussian-flavored distribution driven by the market term, defeating
the purpose of fitting a heavy-tailed generator in the first place.
The dual failure mode, drawing the residual from a Gaussian with
variance taken from the real ordinary least squares (OLS) fit,
recovers the SIM coefficients exactly but erases the marginal
structure entirely. Neither naive option satisfies both targets at
once, and closing that gap is the problem the rest of this paper
addresses.

This paper describes a closed-form variance correction that sits
between the two failure modes. The construction rescales the
per-asset generator draw by a scalar chosen so that the variance
algebra works out to preserve the generator marginal; the scalar
admits a closed form whenever the market component accounts for
less than the full generator variance. When that condition fails
(high-loading assets paired with a low-volatility generator, or a
synthetic market more volatile than the calibration market), the
factor loading is clipped downward against a floor that reserves a
configurable share of the marginal variance for idiosyncratic
noise, with a sign-preserving clip that keeps defensive assets
defensive. A separate branch handles index-tracking assets, where
the calibrated coefficient of determination (the $R^2$ from OLS of
the asset's real growth-rate history against the market path)
rather than the generator marginal is the quantity worth preserving;
SPY against itself recovers as the degenerate perfect-tracking
limit with no special-casing. Branch
selection is driven entirely by the calibrated $R^2$, and each
asset carries a persisted construction flag (\texttt{hybrid},
\texttt{hybrid-clipped}, \texttt{r2-preserve}) so downstream
consumers can audit which path each asset took. The procedure
generalizes to any new ticker without external metadata.

In this study, we tested the construction on the $424$-ticker US
equity universe and SPY market path used by the companion
\texttt{JumpHMM.jl} validation paper of Alswaidan and
Varner~\cite{alswaidanVarner2026jdiq}, comparing three composition
recipes on the same generators and the same market path: the naive
composition, a Gaussian-residual SIM that recovers the calibrated
factor coefficients and $R^2$ but loses the marginal structure, and the
hybrid construction proposed here. We measured SIM
recovery, marginal fidelity, and cross-sectional dependence, reporting
each per construction-flag branch. The hybrid construction held the
SIM-recovery error of the naive composition while holding the
marginal-fidelity error of the Gaussian-residual baseline across the
full range of empirical factor loadings; on the aggregate scorecard
it recovered the factor loading to a median absolute error of $0.018$ and passed a
per-ticker Kolmogorov--Smirnov test against the real marginal at rate
$50.4\%$, compared to $4.5\%$ for the naive composition and $0.6\%$
for the Gaussian baseline. The broader implication is that the
marginal-versus-factor tradeoff long assumed in SIM composition is not
fundamental: a single closed-form scalar correction closes the gap,
and the correction generalizes to richer factor structures
(multi-factor SIM, lagged-factor SIM) without changing the underlying
algebra.